Connectome inferences using partial correlation from imaging calcium signal

### 2014-06-13 article.tex
notation detail + wording

% Introduce notation + goal in mathematical score
% TODO check convention for y_{ij}
The connectomes is represented by a (un)directed graph $G = (V, E)$ comprising
a set of nodes $V$ and a set of edges $E \subseteq \left\{(i, j) \in V \times
V\right\}$, (un)ordered pair from $V$.
Each node $i \in V$ represents a neuron, and each edge $(i, j) \in E$
represents a direct link from neuron $i$ to neuron $j$.
The network inference task is set as follows :
\textit{Given the observations $(x_t \in \mathcal{R}^{|V|})_{t=1}^T $
along $T$ periods of time of the $|V|$ neurons, the goal is to infer the
adjacency matrix $A : a_{i,j} = I((i, j) \in E)$ of the graph $G$,
where $I$ is the indicator function.}

@@

Connectivity information between two neurons may be retrieved from successive
and slightly delayed burst events. Direct links
between two neurons should have shorter delay between fluorescence
peaks than undirected links. However, the fluorescence signal has a slow
decay compare to its rise. Therefore, we filter those unrelevant information
with a hard thresholding filter
\begin{align}
X^{\prime\prime\prime}_{t,i} &=
{X^{\prime\prime}_{t,i}}^{0.9} I(X^{\prime\prime}_{t,i} \geq \tau) \forall i
\label{eq:hard-treshold-filter}
\end{align}
where $\tau \geq 0$ is a parameter of the hard-threshold filter and the $0.9$
exponent reduce the difference between peaks values.

@@

Let us denote the fluorescence calcium indicators of all neuron
as a set of random variables $X = \left\{X_1, \ldots, X_{|V|}\right\}$, which follows
a joint probability distribution $p_\mathcal{X}$, and by
$X^{-i,j}$ the set of all variables except $X_i$ and $X_j$.
Two random variables $X_i$ and $X_j$ are said to be conditionally independent
to a (possibly empty) set of random variables $Z$, denoted by $X_i \perp X_j | Z$,
if the joint conditional probability distribution factorizes
$p_{\mathcal{X}_i, \mathcal{X}_j|\mathcal{Z}} = p_{\mathcal{X}_i|\mathcal{Z}}
p_{\mathcal{X}_j|\mathcal{Z}}$.  The inference of the undirected network
can be formulated as inferring conditional dependences
$X_i \not\perp X_j | X^{-i,j}$ between all pair of fluorescence signal
$i$ and $j$, in other words the inferred network will be
$\hat{A}: \hat{a}_{i,j} = I(X_i \not\perp X_j | X^{-i,j}), \forall i, j \in V$.

@@

If the random variables $X$ follows a joint Gaussian distribution
$p_\mathcal{X} \sim \mathcal{N}(\mu; \Sigma)$ of mean vector $\mu$ and
covariance matrix $\Sigma$, we have the conditional independence
$X_i \perp X_j | X^{-i,j}$ if the partial correlation coefficient
\[
\rho_{X_i, X_j | X^{-i,j}}
= \frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}}.
\]
is equal zero. The partial correlation gives a connectivity score
$\hat{A}: \hat{a}_{i,j} = \rho_{X_i, X_j | X^{-i,j}}, \forall i, j \in V$ between all
pairs of neuron $i$ and $j$. One could retrieve an estimated graph by thresholding
the matrix $\hat{A}$.
By contrast, we also have $X_i \perp X_j$ if the Pearson correlation
coefficient
\[
\rho_{X_i,X_j} = \frac{\Sigma_{ij}}{\sqrt{\Sigma_{ii}
\Sigma_{jj}}}
\]
is equal to zero.

>>>>

% Introduce notation + goal in mathematical score
% TODO check convention for y_{ij}
The connectomes is represented by a (un)directed graph $G = (V, E)$ comprising
a set of nodes $V$ and a set of edges $E \subseteq \left\{(i, j) \in V \times
V\right\}$, (un)ordered pair from $V$.
Each node $i \in V$ represents a neuron, and each edge $(i, j) \in E$
represents a direct link from neuron $i$ to neuron $j$.
The network inference task is set as follows :
\textit{Given the observations $(x_t \in \mathcal{R}^{|V|})_{t=1}^T $
along $T$ periods of time of the $|V|$ neurons, the goal is to infer the
adjacency matrix $A : a_{i,j} = \mathbb{1}((i, j) \in E)$ of the graph $G$,
where $\mathbb{1}$ is the indicator function.}

@@

Connectivity information between two neurons may be retrieved from successive
and slightly delayed burst events. Direct links
between two neurons should have shorter delay between fluorescence
peaks than undirected links. However, the fluorescence signal has a slow
decay compare to its rise. Therefore, we filter those unrelevant information
with a hard thresholding filter
\begin{align}
X^{\prime\prime\prime}_{t,i} &=
{X^{\prime\prime}_{t,i}}^{0.9} \mathbb{1}(X^{\prime\prime}_{t,i} \geq \tau) \forall i
\label{eq:hard-treshold-filter}
\end{align}
where $\tau \geq 0$ is a parameter of the hard-threshold filter and the $0.9$
exponent reduce the difference between peaks values.

@@

Let us denote the fluorescence calcium indicators of all neuron
as a set of random variables $X = \left\{X_1, \ldots, X_{|V|}\right\}$, which follows
a joint probability distribution $p_\mathcal{X}$, and by
$X^{-i,j}$ the set of all variables except $X_i$ and $X_j$.
Two random variables $X_i$ and $X_j$ are said to be conditionally independent
to a (possibly empty) set of random variables $Z$, denoted by $X_i \perp X_j | Z$,
if the joint conditional probability distribution factorizes
$p_{\mathcal{X}_i, \mathcal{X}_j|\mathcal{Z}} = p_{\mathcal{X}_i|\mathcal{Z}}
p_{\mathcal{X}_j|\mathcal{Z}}$.  The inference of the undirected network
can be formulated as inferring conditional dependences
$X_i \not\perp X_j | X^{-i,j}$ between all pair of fluorescence signal
$i$ and $j$, in other words the inferred network will be
$\hat{A}: \hat{a}_{i,j} = \mathbb{1}(X_i \not\perp X_j | X^{-i,j}), \forall i, j \in V$.

@@

If the random variables $X$ follows a joint Gaussian distribution
$p_\mathcal{X} \sim \mathcal{N}(\mu; \Sigma)$ of mean vector $\mu$ and
covariance matrix $\Sigma$, we have the conditional independence
$X_i \perp X_j | X^{-i,j}$ if the partial correlation coefficient
\[
\rho_{X_i, X_j | X^{-i,j}}
= \frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}}
\]
is equal zero. The partial correlation gives a connectivity score
$\hat{A}: \hat{a}_{i,j} = \rho_{X_i, X_j | X^{-i,j}}, \forall i, j \in V$ between all
pairs of neuron $i$ and $j$. One could retrieve an estimated graph by thresholding
the matrix $\hat{A}$.
Similarly, we also have $X_i \perp X_j$ if the Pearson correlation
coefficient
\[
\rho_{X_i,X_j} = \frac{\Sigma_{ij}}{\sqrt{\Sigma_{ii}
\Sigma_{jj}}}
\]
is equal to zero.
